Beta to Release Matching: Model Validation

### 2020-03-17 poc/matchIt/model_validation.Rmd
updating the model validation

df_train_encoder <- read.csv("df_train_encoder.csv", header = T, sep = "\t", encoding="UTF-8") 
df_validate_encoder <- read.csv("df_validate_encoder.csv", header = T, sep = "\t", encoding="UTF-8")
#df_train_encoder$default_search_engine_missing <- 0
```

@@

#### Diagnostics ####
compare_log_cont <- function(df, covariate, tr_mean = NULL, tr_median = NULL, ho_mean = NULL, print = TRUE){
  p <- ggplot(df %>% mutate(!!covariate := get(covariate)+1), aes(label, get(covariate))) +
    geom_violin(aes(fill = label), alpha=0.5) +
    geom_boxplot(width=0.1) +
    scale_y_log10() + scale_fill_manual(values = c("#111d5e", "#b21f66","#611e62")) +
    theme_bw() +
    guides(fill = FALSE) +
    labs(x='Channel', y='Measure', title=covariate) + coord_flip()

@@

```{r include=FALSE}
df_validate_1x$label <- mapvalues(df_validate_1x$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
df_validate_encoder$label <- mapvalues(df_validate_encoder$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
df_train_1x$label <- mapvalues(df_train_1x$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
```

@@

### Match using Nearest Neighbor matching - Mahalanobis: Full Dataset

@@

```{r message=FALSE, warning=FALSE}
nn.ma <- matchit(formula = generate_formula(covariates, label), df_train_1x, 'nearest', distance='mahalanobis')
df_matched <- match.data(nn.ma)
df_matched$label <- mapvalues(df_matched$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
print(summary(nn.ma))
```

```{r}
table_match <- CreateTableOne(vars = covariates, strata = "is_release", data = df_matched, test = FALSE)
print(table_match, smd = TRUE)
```

```{r fig.height=5, fig.width=9, message=FALSE, warning=FALSE}
love.plot(nn.ma, binary = "std", threshold = .1, colors = c("#111d5e", "#b21f66"), abs = T) + theme_bw()
```

***
### Observations

* Both table and plot show that the matching process successfully match all instances. That is, we have no unmatched (unadjusted) samples.
* The plot shows that, for adjusted cases (after matching), the standardized mean difference is relatively small for most covariates. The smallest (that is, less than or equal to $0.1$) are: `profile_age`, `memory_mb`, `num_bookmarks` and `num_active_days`. However, some have a high absolute value, for example, `num_addons`.

### Post-matching Beta-Release Difference:

@@

### QQ-Plots

Q-Q plots are a handy tool for comparing distributions: the closer the values form a straight line, the higher
chance to come from a similar distribution. Here, we use Q-Q plots and the Kolmogorov-Smirnov test (KS) test to verify the differences between the balanced Beta and Release for that version (N).

@@

## QQ plot in R to compare two data samples
for (i in engagement) {

@@

  rg_t <- range(x_t, y_t, na.rm=T)

  test_t <- ks.test(x_t, y_t)$statistic
  test_t <- paste("KS Test = ", round(test_t, 3))

  ########

  title_t <- paste('V67', i, sep='\n')
  qqplot(x_t, y_t, main=title_t, xlim=rg_t, ylim=rg_t, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x_t), (if(max(x_t) > max(y_t)) max(x_t) else max(y_t)), test_t, adj=c(0,1))
  abline(0,1, col="#fe346e", lty=2)
}
```

### Violin Plots

The following violin plots depicts distributions for the Beta and release subsets, for the version v67 (training).

```{r fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
par(mfrow = c(4, 2))

## Violin plots
for (i in engagement) {
  p <- ggplot(df_matched, aes(x=label, y=eval(as.name(i)), fill=label)) + 
  geom_violin(trim=FALSE) +
  labs(title=i,x="Channel", y = "Measure") + 
  scale_y_log10() + scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  geom_boxplot(width=0.1, fill="white", alpha=0.5) + 
  theme_bw() + theme(legend.position="none") + coord_flip()
  print(p)
}
```

### Observations

* Analysing only the Q-Q plots, the best matching variables are near the $x = y$ line (e.g., `active_hours`, `num_pages`, `search_count` and `daily_unique_domains`). However, some plots present a deviation from such a line at the high-end (e.g., `daily_max_tabs` and `daily_tabs_opened`). 
* Regarding violin plots, overall, the Beta users from V67, with all the other covariates (*experiment 3*) being equal, are quite similar concerning Release in user engagement. Only the following metrics yielded a great deviation (as already pointed by the KS test):

  - `num_pages` and `num_pages_max`
  - `daily_max_tabs` and `daily_max_tabs_max`

@@

stats <- calc_stats(df_validate_full, engagement, add_1 = TRUE) %>%
  dplyr::select('metric','label', everything())

@@

## Training Covariates

* Experiment 3:
  - daily_num_sessions_started
  - daily_num_sessions_started_max
  - FX_PAGE_LOAD_MS_2_PARENT
  - memory_mb
  - num_active_days
  - num_addons
  - num_bookmarks
  - profile_age
  - session_length
  - session_length_max
  - TIME_TO_DOM_COMPLETE_MS
  - TIME_TO_DOM_CONTENT_LOADED_END_MS
  - TIME_TO_DOM_INTERACTIVE_MS
  - TIME_TO_LOAD_EVENT_END_MS
  - TIME_TO_NON_BLANK_PAINT_MS

```{r}
stats_pre <- calc_delta(df_validate_1x, covariates)
stats_post <- calc_delta(df_validate_matched, covariates)

```{r fig.width=10, fig.height=50, echo=FALSE, results = 'asis'}
stats <- calc_stats(df_validate_full, covariates, add_1 = TRUE) %>%
  dplyr::select('metric','label', everything())

@@

for (covariate in covariates) {
  stats_rel <- stats %>% filter(label == 'release') %>% dplyr::select(covariate, metric)
  means <- stats_rel[stats_rel$metric == 'mean', covariate]
  medians <- stats_rel[stats_rel$metric == 'median', covariate]
  ho_means <- stats %>% filter(label == 'beta' & metric == 'mean') %>% dplyr::select(covariate)

@@

>>>>

df_train_encoder <- read.csv("df_train_encoder.csv", header = T, sep = "\t", encoding="UTF-8") 
df_validate_encoder <- read.csv("df_validate_encoder.csv", header = T, sep = "\t", encoding="UTF-8")
df_train_encoder$default_search_engine_missing <- 0
```

@@

#### Diagnostics ####
compare_log_cont <- function(df, covariate, tr_mean = NULL, tr_median = NULL, ho_mean = NULL, print = TRUE){
    p <- ggplot(df %>% mutate(!!covariate := get(covariate)+1), aes(label, get(covariate))) +
        geom_violin(aes(fill = label), alpha = 0.5) +
        geom_boxplot(width=0.1) +
        scale_y_log10() + scale_fill_manual(values = c("#111d5e", "#b21f66","#611e62")) +
        theme_bw() +
        guides(fill = FALSE) +
        labs(x = 'Channel', y='Measure', title=covariate) + coord_flip()

    # add in lines for treatment means/medians
    if (!is.null(tr_mean)){
        p <- p + 
            geom_hline(yintercept=tr_mean, linetype="solid", lwd = .5) + 
            geom_hline(yintercept=tr_median, linetype="dashed", lwd = .5) + 
            geom_hline(yintercept=ho_mean, linetype="dashed", color = "#D62728B2", lwd = 1)
    }

@@

```{r include=FALSE}
df_validate_1x$label <- mapvalues(df_validate_1x$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
df_validate_encoder$label <- mapvalues(df_validate_encoder$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
df_train_1x$label <- mapvalues(df_train_1x$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
df_train_encoder$label <- mapvalues(df_train_encoder$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
```

@@

### Match using Nearest Neighbor matching: Full Dataset

@@

```{r message=FALSE, warning=FALSE}
nn <- matchit(formula = generate_formula(covariates, label), df_train_1x, 'nearest', replace = TRUE)
df_matched <- match.data(nn)
df_matched$label <- mapvalues(df_matched$is_release, from=c(FALSE,TRUE), to=c('beta','release'))
print(summary(nn))
```

```{r}
table_match <- CreateTableOne(vars = covariates, strata = "label", data = df_matched, test = FALSE)
print(table_match, smd = TRUE)
```

```{r fig.height=5, fig.width=9, message=FALSE, warning=FALSE}
love.plot(nn, binary = "std", threshold = .1, shapes = c("triangle filled", "circle filled"), colors = c("#111d5e", "#b21f66"), abs = F, position = "top", var.order = "unadjusted") + theme_bw()
```

***
#### Observations

* The Love plot is a summary plot of covariate balance pre and post matching. Each point represents the balance statistic for that covariate, colored based on whether it is calculated before or after adjustment. The dotted lines represent the threshold set ($0.1$); if most or all of the points after adjustment are within the threshold, that is good evidence that balance has been achieved
* The plot shows that, for unadjusted cases (pre-matching), the standardized mean difference is not relatively large (except for `num_addons`). However, for adjusted cases (post-matching), the standardized mean difference is smaller. That is, for most cases, the absolute value is even smaller than the threshold ($0.1$)
* Therefore, this result is a good evidence that the matching worked well, as the balance was improved on almost all variables after adjustment

### Post-matching Beta-Release Difference

@@

```{r fig.width=10, fig.height=50, echo=FALSE, results = 'asis'}
df_training_full <- df_matched %>% 
  dplyr::select(-weights, -distance) %>%
  filter(label == 'beta') %>%
  mutate(label =  'beta - matched') %>%
  rbind(df_train_1x) %>% 
  distinct()

stats <- calc_stats(df_training_full, engagement) %>%
  dplyr::select('metric','label', everything())

### Kolmogorov-Smirnov test (KS)

We can use the Kolmogorov-Smirnov test (KS) to verify the differences between the balanced Beta and Release for the v67 version. It is between 0 and 1, and represents how two data sets are similar. Smaller KS distance values indicate better balance.

@@

output = data.frame()

for (i in engagement) {

@@

  rg_t <- range(x_t, y_t, na.rm=T)
  ks <- ks.test(x_t, y_t)$statistic
  output = rbind(output, data.frame(KS = ks))
}
```

```{r message=FALSE, warning=FALSE, echo=FALSE, results = 'asis'}
kable(set_rownames(output, engagement)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

### Visual inspection

Here, we display density plots for the two groups on the given user engagement metric, so we can visually compare their distribution. The degree to which the densities for the two groups overlap is a good measure of group balance on the given covariate; significant differences in shape can be indicative of poor balance, even when the mean differences and variance ratios are well within thresholds.

```{r echo=FALSE, fig.height=4, fig.width=7}
par(mfrow = c(7, 2))
for (v in engagement){
  p <- ggplot(data=df_matched, aes(x = get(v), group = label, fill = label)) +
    geom_density(aes(fill = label), adjust=1.5, alpha=.4) + 
    scale_fill_manual(values = c("#111d5e", "#b21f66")) +
    labs(y='Density', title=v) + theme_bw() + theme(legend.position="bottom") 
  print(p)
}
```

The following violin plots depicts distributions for the following subsets:

* Beta v67: pre-matching
* Beta v67: matched and subsetted
* Release v67

**NOTE**: Guiding lines have been added for the following:

* <span style="color:black">black</span> solid: Release mean
* <span style="color:black">black</span> dashed: Release median
* <span style="color:#D62728B2">red</span> dashed line: subsetted Beta _mean_.

  plots[[covariate]] <- compare_log_cont(df_training_full, covariate, means, medians, as.numeric(ho_means), print=FALSE) 
}
```

***
#### Observations

The density and violin plots show that there are significant differences between both groups (Beta and Release) concerning some user engagement metrics, listed as follows.

* `num_pages`
* `num_pages_max`
* `active_hours`
* `active_hours_max`
* `uri_count`
* `uri_count_max`
* `daily_unique_domains`
* `daily_unique_domains_max`

@@

stats <- calc_stats(df_validate_full, engagement) %>%
  dplyr::select('metric','label', everything())

@@

### Kolmogorov-Smirnov test (KS)

Once again, we use the KS test to verify whether any significant difference between the average user engagement metrics in the Beta and Release groups, over several versions (v67 and v68). *Reminder:* smaller KS distance values indicate a better balance.

df_beta <- df_validate_full %>% filter(label == 'beta - matched')
df_rel <- df_validate_full %>% filter(label == 'release')

output = data.frame()

for (i in engagement) {

  rg_t <- range(x_t, y_t, na.rm=T)
  ks <- ks.test(x_t, y_t)$statistic
  output = rbind(output, data.frame(KS = ks))
}
```

```{r message=FALSE, warning=FALSE, echo=FALSE, results = 'asis'}
kable(set_rownames(output, engagement)) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

### Visual inspection

```{r echo=FALSE, fig.height=4, fig.width=7}
par(mfrow = c(7, 2))
for (v in engagement){
  p <- ggplot(data=df_validate_full, aes(x = get(v), group = label, fill = label)) +
    geom_density(aes(fill = label), adjust=1.5, alpha=.4) + 
    scale_fill_manual(values = c("#111d5e", "#b21f66","#611e62")) +
    labs(y='Density', title=v) + theme_bw() + theme(legend.position="bottom") 
  print(p)
}
```

The following violin plots depicts distributions for the following subsets:

* Beta v68: pre-matching
* Beta v68: matched and subsetted
* Release v68

**NOTE**: Guiding lines have been added for the following:

* <span style="color:black">black</span> solid: Release mean
* <span style="color:black">black</span> dashed: Release median
* <span style="color:#D62728B2">red</span> dashed line: subsetted Beta _mean_.

@@

@@

***
#### Observations

Our main objective was to determine if the user engagement metrics changed in the newest Beta version concerning the previous Release version. The density and violin plots show that there are significant differences between both groups (Beta and Release) concerning some user engagement metrics, listed as follows.

* `daily_max_tabs`
* `daily_max_tabs_max`
* `num_pages`
* `num_pages_max`
* `daily_unique_domains`

```{r eval=FALSE, include=FALSE}
# Saving
save.image(file = "model_validation.RData")
```
